---
title: "Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(interactions)
library(nlme)
library(performance)
library(tidyr)
library(jtools)
library(broom)

library(pacman)

# These are the packages you will need for the analyses 
p_load(here, tidyverse, DT, ggplot2, xaringan, knitr, kableExtra, modelsummary, stargazer, xaringanthemer, ggthemes, fixest, haven, arsenal)
data_owid <- readRDS("data_owid.RDS")

```


# Unconditional model, calculate ICC
```{r m0}
# random effect. Intercept varies by clusters (name)
# Deafault: t-tests use Satterthwaite's method . For small sample, use ddf = "Kenward-Roger" to adjusted small-sample bias. 
m0 <- lmer(total_deaths_per_million ~ (1|name), 
                     data = data_owid, 
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))

summary(m0)
#Intercept in a model w/o predictor means the grand mean of outcome. 
icc(m0)
```
The fixed effect for the intercept means that the grand mean of Covid mortality rate per million population is 472.48 for the pooled data (all time point months in all countries). 

The level-1 residual variability is 299012, the level-2 random intercept variance is 291718. The intra-class correlation ICC is 291718/(291718 + 299012) =  .494, which means the model implies that 49.4% of the variance in Covid mortality is associated with between-country differences. Equivalently, on average, within a given country, Covid mortality rate has a .494 correlation.  

# Linear Growth Model 
This model includes both random intercepts (1) and random slopes (time_month). 

```{r m_timelinear}
m_timelinear <- lmer(total_deaths_per_million ~ time_month + (1 + time_month|name), 
                     data = data_owid, 
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
summary(m_timelinear)
#confint(m_timelinear)
```
Time was a significant predictor of mortality rate. 
Compared to the unconditional (empty) model above, the variance associated with name has been reduced from 291718 to 40218.

Variance among the intercepts: 40218
Variance among the slopes:3297
Residual variance was substantially reduce from 299012 in the previous model to 20950 in this model that added time as a predictor. 

Correlation between slope and intercept: -0.53

The model's result shows that the initial average value .  

# Quadratic Unconditional Growth Model 

```{r}
m_timesq <- lmer(total_deaths_per_million ~ time_month + I(time_month^2) + (time_month|name),
                 data = data_owid, 
                 control = lmerControl(optimizer = "bobyqa", 
                                       optCtrl = list(maxfun = 2e5)))
summary(m_timesq)
```

Test improvement in model fit with quadratic term. 

##Visualize model fit
https://mlm2-2021.netlify.app/slides/w6p1.html#114
```{r }
# NOTE: need to set REML = FALSE, because with restricted maximum likelihood (REML) estimation, likelihood ratio test can only be conducted between models that are nested in the covariance parameters. The nesting relationship includes the fixed effect, so can NOT use a likelihood ratio test to compare the models when fit by REML. --> need full-information maximum likelihood FIML.  

m_timelinear_FIML <- lme4::lmer(total_deaths_per_million ~ time_month + (time_month|name), 
                     REML = FALSE,
                     data = data_owid, 
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))

m_timesq_FIML <- lme4::lmer(total_deaths_per_million ~ time_month + I(time_month^2) + (time_month|name),
                  REML = FALSE,
                 data = data_owid, 
                 control = lmerControl(optimizer = "bobyqa", 
                                       optCtrl = list(maxfun = 2e5)))
anova(m_timelinear_FIML, m_timesq_FIML)
# result shows m_timesq is a big improvement of model fit. 
```


# Testing for Heteroscedastic Errors --> Need to account for heteroscedasticity. 

```{r}
# to override the default homoscedastic error structure for level 1 residuals, need to use nlme; Note that lme4 does not allow for testing of the homoscedasticity assumption. 
m_test_homo2 <- nlme::lme(total_deaths_per_million ~ time_month + 1,
          data = data_owid, 
          random = ~ time_month + 1|name,   # random effect of time and intercept 1 --> dependent on name
          control = lmeControl(opt='optim'), # to avoid LME() error - iteration limit reached
          weights = varIdent(form =~ 1|time_month)) # allow for heteroscedastic error variance over time --> weights indicating that the residual variances is a function of time

summary(m_test_homo2)
intervals(m_test_homo2)
```


# System of Internal Governance: Unitary, Semi-federal, Federal states  

```{r}
# list of federal
# including semi-federal
data_owid %>% 
  filter(unitary_semi_federal == 1) %>% distinct(name) -> f1

# NOT including semi-federal
data_owid %>% 
  filter(unitary_federal == 1) %>% distinct(name) -> f2

# differences between 2 lists
setdiff(f1, f2) -> f3
f3
```

## Method 1: 22 federal states --> not significant

```{r}
m_unitary_federal <- lmer(total_deaths_per_million ~ time_month + I(time_month^2) +  (time_month|name) +                                  log(gdp_percap) + health_index +
                            time_month + unitary_federal, 
                          data = data_owid, 
                          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_unitary_federal)
```

## Method 2: 32 federal states --> Significant

In this model, the form of state organization - federal vs. unitary states - is a categorical predictor at level 2 (country level). 

```{r m_unitary_federal}
m_unitary_semi_federal <- lmer(total_deaths_per_million ~ time_month + I(time_month^2) +  (time_month|name) + 
                             log(gdp_percap) + health_index + 
                              unitary_semi_federal, 
                          data = data_owid, 
                          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m_unitary_semi_federal)
```

Considering the fixed effect results, the main effect of internal governance system (federal coded 1) is significant, p = 0.016. The model implies that, on average, compared to a state under a unitary system, a federal state has intercept higher than unitary state by 92.8 Covid death per million population, and this significantly differs from zero (t = 2.42, SE = 38.2, p = 0.016). 

NOTE: when including the cross-level interaction between time and federal represents the regression of trajectory slopes on federal is not significant (gamma = , p = ). This means that the model-implied slopes do not differ in magnitude of slope between federal and unitary states. 

```{r sim_slope, unitary_federal}
 sim_slopes(m_unitary_semi_federal, pred = time_month, modx = unitary_semi_federal, 
            data = data_owid,
            centered = "none", cond.int=TRUE,
  johnson_neyman = FALSE)
```


```{r interact_plot, unitary_federal}
interact_plot(m_unitary_semi_federal, pred = time_month, modx = unitary_semi_federal,
              data = data_owid,
 x.label = "Time point (in month)",
 y.label = "Mortality",
 main.title = "Covid Mortality per million population \n Federal vs. Unitary system ",
 legend.main = "")

```
Significant differences in intercepts, not in slopes. 

Both federal models 
```{r}
modelsummary(list(m_unitary_federal, m_unitary_semi_federal), 
               title = "Table  ",
               stars=c('*' = 0.05, '**' = 0.01, '***' = 0.001),
               coef_omit = "(Intercept)|as.factor",
               #coef_rename = c("health_index" = "Health Security Index"),
               estimate = "{estimate}{stars}",
               gof_omit= "Adj|Log|Within|AIC|BIC|FE|Std|F",
               #add_rows = row,
               threeparttable= T,
               notes = c("* p < 0.05, ** p < 0.01, *** p < 0.001 Notes: Standard errors shown in parentheses."),
               output = "federal_modelstable.docx")
```


# Democracy
```{r}
m_join_demo <- lmer(total_deaths_per_million ~ I(time_month^2) + (time_month|name) + 
                       log(gdp_percap) + health_index +
                        unitary_semi_federal +
          time_month*scale(democracy), # when have interaction term, lower-order term is included by default. 
                     data = data_owid, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_join_demo)
```

```{r sim_slope, interaction plot, democracy}
 sim_slopes(m_join_demo, pred = time_month, modx = democracy, 
            data = data_owid,
            centered = "none", cond.int=TRUE,
  johnson_neyman = FALSE)

interact_plot(m_join_demo, pred = time_month, modx = democracy,
              data = data_owid,
 x.label = "Time point (in month)",
 y.label = "Mortality",
 main.title = " ",
 legend.main = "")

```
Accordingly, countries with high democracy started out with lower number of Covid death per million population, but its trend increases quicker. Countries with lower democracy started out at more death but its trend increases slower. In other words, higher democratic countries had less death early in the pandemic. Starting at around time point 3 (July 2020), they suffered more from Covid death compared to lower democratic countries.

# Gove_eff 
## Random slope, main effect
Let the slope of Covid mortality to vary by country

```{r}
m_join5 <- lmer(total_deaths_per_million ~ time_month + I(time_month^2) + (time_month|name)  +
                  log(gdp_percap) + health_index + 
           unitary_semi_federal +
           time_month*scale(democracy) +
           scale(gov_eff),
          data = data_owid, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m_join5)
```
## random slope, interaction with slope (time)
```{r}
m_join6 <- lmer(total_deaths_per_million ~ time_month + I(time_month^2) + (time_month|name)  +
                  log(gdp_percap) + health_index + 
           unitary_semi_federal +
           time_month*scale(democracy) +
           time_month*scale(gov_eff),
          data = data_owid, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m_join6)
```


# Trust

In a smaller sample size, both democracy and trust in government have significant negative effect on the slope of Covid mortality. 
```{r}
m_join_trust <- lmer(total_deaths_per_million ~ time_month + I(time_month^2) + (time_month|name) + 
                       log(gdp_percap) + health_index +
                        unitary_semi_federal +
                        time_month*scale(democracy) +
                        scale(gov_eff) + 
                        time_month*scale(trust_gov),
                     data = data_owid, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_join_trust)

```
In the join model with all 4 factors, democracy behave differently. High democracy started high, but get flatter slope. Low democracy started higher, then increase faster. 

```{r interact_plot for democracy, small sample size}
sim_slopes(m_join_trust, pred = time_month, modx = democracy, 
            data = data_owid,
            centered = "none", cond.int=TRUE,
  johnson_neyman = FALSE)

interact_plot(m_join_trust, pred = time_month, modx = democracy,
              data = data_owid,
 x.label = "Time point (in month)",
 y.label = "Mortality",
 main.title = " ",
 legend.main = "")

```


# Create table

```{r}
modelsummary(m_unitary_semi_federal, 
               title = "Table 1. ",
               stars=c('*' = 0.05, '**' = 0.01, '***' = 0.001),
               coef_omit = "(Intercept)|as.factor",
               coef_rename = c("health_index" = "Health Security Index", 
                               "unitary_semi_federal" = "Federal system (Yes = 1)",
                               
                               "log(gdp_percap)" = "GDP per capita (log)",
                               "time" = "Time point (quarter)"
                               ),
               estimate = "{estimate}{stars}",
               gof_omit= "Adj|Log|Within|AIC|BIC|FE|Std|F",
               #add_rows = row,
               threeparttable= T,
               notes = c("* p < 0.05, ** p < 0.01, *** p < 0.001 Notes: Standard errors shown in parentheses."),
               output = "model1.docx")
```


```{r}

# Produce the table; can export to markdown, tex, etc. by changing the type
modelsummary(list(m_unitary_semi_federal, m_join_demo, m_join5, m_join_trust), 
               title = "Table 1. ",
               stars=c('*' = 0.05, '**' = 0.01, '***' = 0.001),
               coef_omit = "(Intercept)|as.factor",
               coef_rename = c("health_index" = "Health Security Index", 
                               "unitary_semi_federal" = "Federal system (Yes = 1)",
                               "democracy" = "Democracy",
                               "trust_gov" = "Trust in government",
                               "gov_eff" = "Government Effectiveness",
                               "log(gdp_percap)" = "GDP per capita (log)",
                               "time" = "Time point (quarter)"
                               ),
               estimate = "{estimate}{stars}",
               gof_omit= "Adj|Log|Within|AIC|BIC|FE|Std|F",
               #add_rows = row,
               threeparttable= T,
               notes = c("* p < 0.05, ** p < 0.01, *** p < 0.001 Notes: Standard errors shown in parentheses."),
               output = "modelstable.docx")

```


