---
title: "Models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(lmerTest)
library(interactions)
library(nlme)
library(performance)
library(tidyr)
library(jtools)
library(broom)

library(pacman)

# These are the packages you will need for the analyses 
p_load(here, tidyverse, DT, ggplot2, xaringan, knitr, kableExtra, modelsummary, stargazer, xaringanthemer, ggthemes, fixest, haven, arsenal)
data <- readRDS("data.RDS")

```


# Unconditional, ICC
```{r m0}
m0 <- lmer(death_mean_quarter ~ (1|name), 
                     data = data, 
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))

summary(m0)

icc(m0)
```

The model implies that 46.2% of the variance in mortality is associated with between-country differences. In other words, within a given country, mortality rate have a .462 correlation. 

## 1.2 Linear Growth Model 
Time was a significant predictor of mortality rate 
```{r m_timelinear}
m_timelinear <- lmer(death_mean_quarter ~ time + (1 + time|name), 
                     data = data, 
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))
summary(m_timelinear)
confint(m_timelinear)
```


Variance among the intercepts: 
Variance among the slopes:
Level 1 residual variance: 

The model's result shows that the initial average value .  

## 1.3 Testing nonlinear functions of time - Quadratic Unconditional Growth Model 

```{r m_timesq}
m_timesq <- lme4::lmer(death_mean_quarter ~ time + I(time^2) + 
                   (1 + I(time^2)|name),
                
                 data = data, 
                 control = lmerControl(optimizer = "bobyqa", 
                                       optCtrl = list(maxfun = 2e5)))
summary(m_timesq)
```

Test improvement in model fit with quadratic term. 

Visualize model fit
https://mlm2-2021.netlify.app/slides/w6p1.html#114
```{r }
# NOTE: need to set REML = FALSE, because with restricted maximum likelihood (REML) estimation, likelihood ratio test can only be conducted between models that are nested in the covariance parameters. The nesting relationship includes the fixed effect, so can NOT use a likelihood ratio test to compare the models when fit by REML. --> need full-information maximum likelihood FIML.  

m_timelinear_FIML <- lme4::lmer(death_mean_quarter ~ time + (1 + time|name), 
                     REML = FALSE,
                     data = data, 
                     control = lmerControl(optimizer = "bobyqa", 
                                           optCtrl = list(maxfun = 2e5)))

m_timesq_FIML <- lme4::lmer(death_mean_quarter ~ time + I(time^2) + 
                   (1 + time + I(time^2)|name),
                  REML = FALSE,
                 data = data, 
                 control = lmerControl(optimizer = "bobyqa", 
                                       optCtrl = list(maxfun = 2e5)))
anova(m_timelinear_FIML, m_timesq_FIML)
# result shows m_timesq is a big improvement of model fit. 
```
## Testing for Heteroscedastic Errors --> Need to account for heteroscedasticity. 

```{r}
# to override the default homoscedastic error structure for level 1 residuals, need to use nlme; Note that lme4 does not allow for testing of the homoscedasticity assumption. 

m_test_homo1 <- nlme::lme(death_mean_quarter ~ time + 1,
          data = data, 
          random = ~ time + 1|name) # random effect of time and intercept 1 --> dependent on name
summary(m_test_homo1)
intervals(m_test_homo1)
```

```{r}
m_test_homo2 <- nlme::lme(death_mean_quarter ~ time + 1,
          data = data, 
          random = ~ time + 1|name,   # random effect of time and intercept 1 --> dependent on name
          control = lmeControl(opt='optim'), # to avoid LME() error - iteration limit reached
          weights = varIdent(form =~ 1|time)) # allow for heteroscedastic error variance over time --> weights indicating that the residual variances is a function of time
          
          
summary(m_test_homo2)
intervals(m_test_homo2)
```

```{r}
anova(m_test_homo1, m_test_homo2)
```

# 2. Conditional Growth Model

The following time-invariant predictors were used to fit the conditional growth model:

* Governmental system, federal vs. unitary states: binary variable. Federal states are coded 1. Unitary states are coded 0. 

* Level of democracy: continuous variable. 

* Level of trust in government

* Level of individual liberty


## 2.2 System of Internal Governance: Unitary, Semi-federal, Federal states  


### Method 1: 25 federal states --> not significant

```{r}
m_unitary_federal <- lmer(death_mean_quarter ~ time + I(time^2) +  (time|name) + 
                            time*unitary_federal + log(gdp_percap), 
                          data = data, 
                          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_unitary_federal)
```

```{r }
# sim_slopes(m_unitary_federal, pred = time, modx = unitary_federal, centered = "none", cond.int=TRUE,
#  johnson_neyman = FALSE)
```

```{r}
# interact_plot(m_unitary_federal, pred = time, modx = unitary_federal,
#  x.label = "Time point", 
#  y.label = "Average mortality (quarterly)",
#  main.title = "Rate of Mortality per 100 thousand population \n Federalvs. Unitary system (26 Federal States)",
#  legend.main = "")
```

### Method 2: 36 federal states --> Significant

In this model, the form of state organization - federal vs. unitary states - is a categorical predictor at level 2 (country level). 

```{r m_unitary_federal}
m_unitary_semi_federal <- lmer(death_mean_quarter ~ time + I(time^2) + 
                   (time|name) +

                                 #(1 + time|name) + I(time^2) +
                                 time*unitary_semi_federal + log(gdp_percap), 
                          data = data, 
                          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_unitary_semi_federal)

```



```{r sim_slope, unitary_federal}
 sim_slopes(m_unitary_semi_federal, pred = time, modx = unitary_semi_federal, centered = "none", cond.int=TRUE,
  johnson_neyman = FALSE)
```


```{r interact_plot, unitary_federal}
interact_plot(m_unitary_semi_federal, pred = time, modx = unitary_semi_federal,
 x.label = "Time point",
 y.label = "Average mortality (quarterly)",
 main.title = "Covid Mortality per 100 thousand population \n Federal vs. Unitary system (36 Federal States)",
 legend.main = "")

```


Both federal models 
```{r}
modelsummary(list(m_unitary_federal, m_unitary_semi_federal), 
               title = "Table  ",
               stars=c('*' = 0.05, '**' = 0.01, '***' = 0.001),
               coef_omit = "(Intercept)|as.factor",
               #coef_rename = c("health_index" = "Health Security Index"),
               estimate = "{estimate}{stars}",
               gof_omit= "Adj|Log|Within|AIC|BIC|FE|Std|F",
               #add_rows = row,
               threeparttable= T,
               notes = c("* p < 0.05, ** p < 0.01, *** p < 0.001 Notes: Standard errors shown in parentheses."),
               output = "federal_modelstable.docx")
```


# Trust model 



```{r}
trust_data <- data %>% drop_na(trust_gov) 

length(unique(trust_data$name))

m_trust <- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) +
                  gdp_percap +
            time*trust_gov, 
          data = data, 
          control = lmerControl(optimizer = "bobyqa", 
                                optCtrl = list(maxfun = 2e5)))

summary(m_trust)
```



# 3 Join model



```{r, results='asis'}
library(huxtable)
library(broomExtra)
#https://cran.r-project.org/web/packages/jtools/vignettes/summ.html#summ

# to compare result of several models
# export_summs(m_join0, 
#              to.file = "docx", file.name = "m_join0.docx")
```

### Federal system
```{r join model, sq time}
# unitary vs federal
m_join_feda <- lmer(death_mean_quarter ~  time + I(time^2) + (time|name) + 
                      log(gdp_percap) + # control for gdp
            time*unitary_semi_federal,
            
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_join_feda)
# ==> unitary_semi_federal significant 
m_join_fedb <- lmer(death_mean_quarter ~  time + I(time^2) + (time|name) +
                      log(gdp_percap) + 
                      
            time*unitary_semi_federal +
            time*health_index,  
            
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_join_fedb)
```

### Democracy
```{r}

m_join_demoa <- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) + log(gdp_percap) + # control for gdp
            time*democracy,
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

m_join_demob<- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) + log(gdp_percap) +
             time*democracy +
            time*health_index,
            
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m_join_demoa)
summary(m_join_demob)

```

### Trust

```{r}
m_join_trusta <- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) + log(gdp_percap) +
            time*trust_gov, 
            
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

m_join_trustb <- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) + log(gdp_percap) +
            time*trust_gov +
            time*health_index,
            
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

summary(m_join_trusta)
summary(m_join_trustb)
```



```{r}

m_join4 <- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) + log(gdp_percap) +
            time*unitary_semi_federal +
            time*health_index +    
            time*democracy +
            time*trust_gov, 
            
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))

m_join5 <- lmer(death_mean_quarter ~ time + I(time^2) + (time|name) + log(gdp_percap) + 
            time*unitary_semi_federal +
            time*health_index +    
            time*democracy +
            time*trust_gov + 
            time*gov_eff,
          data = data, 
          control = lmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 2e5)))
summary(m_join1) 
summary(m_join1_b) 
summary(m_join2) 
summary(m_join3) 
summary(m_join4)
summary(m_join5)


```


# Create table

```{r}

# Produce the table; can export to markdown, tex, etc. by changing the type
modelsummary(list(m_timesq, m_join1, m_join2, m_join3, m_join4, m_join5), 
               title = "Table 1. ",
               stars=c('*' = 0.05, '**' = 0.01, '***' = 0.001),
               coef_omit = "(Intercept)|as.factor",
               coef_rename = c("health_index" = "Health Security Index", 
                               "unitary_semi_federal" = "Federal system (Yes = 1)",
                               "democracy" = "Democracy",
                               "trust_gov" = "Trust in government",
                               "gov_eff" = "Government Effectiveness",
                               "log(gdp_percap)" = "GDP per capita (log)",
                               "time" = "Time point (quarter)"
                               ),
               estimate = "{estimate}{stars}",
               gof_omit= "Adj|Log|Within|AIC|BIC|FE|Std|F",
               #add_rows = row,
               threeparttable= T,
               notes = c("* p < 0.05, ** p < 0.01, *** p < 0.001 Notes: Standard errors shown in parentheses."),
               output = "modelstable.html")

```

# Structure Error

https://bookdown.org/roback/bookdown-BeyondMLR/ch-lon.html#learning-objectives-8

# 3. Other considerations (to be continued)

## 3.1 Scaling inputs

Standardizing the inputs does not change model result. 
```{r data_scale, include=FALSE}
data_scale <- data %>%

  select(-trust_gov) %>% drop_na() %>%
  mutate(gdp_percap = scale(gdp_percap),
         health_index = scale(health_index),
         control_corruption = scale(control_corruption),
         gov_eff = scale(gov_eff),
         democracy = scale(democracy),
         civil_liberty = scale(civil_liberty) 
         )


m_scale_inputs <- lmer(death_mean_quarter ~ I(time^2) + (1 + time|name) +
               
            time*unitary_semi_federal +
            time*gdp_percap +
            time*health_index +    

            time*democracy +
            time*civil_liberty +
              
            time*gov_eff + 
            time*control_corruption,
            #time*fragility,
            
          data = data_scale, 
          control = lmerControl(optimizer = "bobyqa", 
                                optCtrl = list(maxfun = 2e5)))
summary(m_scale_inputs) 
```

```{r m_scale_inputs, include=FALSE}
# m_scale_inputs <- lmer(death_mean_quarter ~ time*gov_eff + 
#             time*health_index + 
#             time*democracy +
#             time*civil_liberty +
#             time*gov_index +
#             time*unitary_federal +
#             (1 + time|name), 
#           data = data_scale, 
#           control = lmerControl(optimizer = "bobyqa", 
#                                 optCtrl = list(maxfun = 2e5)))
# summary(m_scale_inputs) 
```



